---
title: "Assignment 10"
author: "Durga Pokharel"
date: "18/01/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
School/University: School of Mathematical Science, Tribhuvan University, Nepal
Roll.no: '11'
Cource Name: Statistical Computing with R(MDS 503)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#[Q.N.1] Create two quantitative variables: age and body-mass index (BMI) with random samples of size 1000 each:
     * Age: 0 to 99 (random samples)
     * BMI: 10 to 40 (random samples)
     
`Age`

```{r cars}
age <- sample(0:99,size = 1000, replace =T)


```

`BMI`
```{r}
bmi <- sample(10:40, size = 1000, replace = T)
```

# [Q.N.2] Create a binary variable sex (1=Male and 0=Female) of 1000 random samples


```{r pressure, echo=FALSE}
sex <- rep(c(0,1),1000,repace=T)



```
# [Q.N.3] Create a data frame as df containing four variables/features:Serial Number, BMI, Age and Sex

```{r}
sn <- c(1:1000)
df <- data.frame(sn,bmi,age,sex)
head(df)
```

# [Q.N.4] For replication of the results, use your class roll number as random.seed during analysis
# [Q.N.5] Split the data into “train” and “test” data using 80-20 partition
```{r}
set.seed(11)
nd = sample(2,nrow(df), replace= T,prob = c(0.8,0.2))
train_data <- df[nd ==1,]
test_data <- df[nd==2, ]
```

# [Q.N,6] Fit a linear regression model with BMI as dependent variable and age and sex and predictors in the train data samples

```{r}
lm1 <- lm(bmi~.,data = train_data)
summary(lm1)
```
Here p value grater than 0.05. Hence dependent variable normailly distributed. 

# [Q.N.7] Conduct residual analysis of the fitted model with graphs (suggestive) and tests (confirmative)

`Linearity of resudial`
**Graphical (Suggestive)**
```{r}
plot(lm1, which = 1, col= c("blue"))
```

**Calculation (Confirmative)** 
```{r}
summary(lm1$residuals)
```

`Independence`
**Graphical (suggestive)**
```{r}
acf(lm1$residuals)
```

**Calculatin(Confirmative)**
```{r,message=FALSE}
library(car)
durbinWatsonTest(lm1)
```

`Normality of Residuals`
**Graphical(suggestive)**
```{r}
plot(lm1, which = 2, col = c("blue"))
```
**Calculation(Confirmative)**
```{r}
shapiro.test(lm1$residuals)
```


`Equal Variance `
**Graphical(Suggestive)**
```{r}
plot(lm1,which = 3, col= c("blue"))
```


**Calculation(Confirmative)**
```{r, message=FALSE}
library(lmtest)
bptest(lm1)
```
# [Q.N.8] Use the fitted model to predict the random test data samples
```{r}
pred <- predict(lm1, data = test_data)
length(pred)
```


```{r}
length(test_data$bmi)
```

# [Q.N.9] Get R-square, MSE and RMSE for training as well as test data and interpret them carefully


```{r, message=FALSE}
library('caret')
data.frame(R2 = R2(pred, 
train_data$bmi),
RMSE = RMSE(pred, 
train_data$bmi),
MAE = MAE(pred, 
train_data$bmi))



```


From this value of r-square only 0.45% data fit the regression model. Which is actually low. Similarly values for RMSE and MAE is respectively 8.909007 and 7.745684. It is better to have low error. Also, I have tried to get value of R2, RMSE and MAE for test data I got error `Error in cor(obs, pred, use = ifelse(na.rm, "complete.obs", "everything")) : 
incompatible dimensions` I tried to solve but I could not able.





