---
title: "Untitled"
author: "Durga Pokharel"
date: "02/12/2021"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#ggplot2

## Grammer :

A grammar provides a foundation for understanding a diffrent type of graphics. A grammar may also help guide us on what a well-formed or correct 
graphic looks like, but there will still be many grammatically correct but nonsensical graphics. This is easy to see by analogy to the English language: good grammar 
is just the first step in creating a good sentence.


## Grammar of graphics:
A grammar of graphics is a tool that from which it enables us to clearly describe the components of a graphic. Such a grammar allows us to move beyond named graphics (e.g., the 
“scatterplot”) and gain insight into the deep structure that underlies statistical graphics. • ggplot2 proposes an alternative parameterization of the grammar, 
based around the idea of building up a graphic from multiple layers of data.

## Components of ggplot2:
* Data and aesthetic mapping +
* Geometric objects
* Scale
* Facet Specification
* Statistical Transformation
* Coordinate Syatem

## Layered grammar of graphics:
Together, the data, mappings,statistical transformation, and geometric object form a layer. Plot may have different layer.Layers are responsible for creating the objects that we expected on 
the plot.

## How to use ggplot2 in R.
For this we should have to installed ggplot2 package in our IDE. Let us use ggplot in R builtin data `diamonds`.
```{r}
library(ggplot2)
ggplot(diamonds, aes(carat,price)) + geom_point()

```
geom_point is used for scatter plot. From above figure we can see that as diamond's carat increases prices is also increases. We can not see how the data distibuted for this let make some changes in our code.

```{r}
ggplot(diamonds,aes(carat,price)) + geom_point() +
  scale_x_continuous() + scale_y_continuous()
```
We can see better distribution of points than previous plot. We clearly see that caret and price variable are not linearly distributed. To make it linealy distributable also make some changes in our code.

```{r}
ggplot(diamonds, aes(carat,price)) + geom_point() +
  stat_smooth(method = lm) + scale_x_log10() + scale_y_log10()
```
From above graph relationship between price and carat variables is linear.
If we try the code without `stat_smooth(method= lm)` we can not see linear line in graph.

```{r}
ggplot(diamonds, aes(carat,price)) + geom_point() +
  scale_x_log10() + scale_y_log10()
```
## Lets make histogram of `diamonds` data.

```{r}
ggplot(diamonds, aes(price)) + geom_histogram()
```
To build histogram we use function `geom_histogram()`. We should note histogram is made on one dimensional data. If we want to add title of the plot we can do,

```{r}
ggplot(diamonds, aes(price)) + geom_histogram() + ggtitle("ggplot2 Histogram")
```
## Let us try some others ggplot2 features in R builtin data `mtcars`.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
geom_point()
```
The figure above scatter plot between 'hwy` and `displ` variables of mtcars data from figure we can see as the values of `hwy` increase vales of `displ` variables slightly decreases.

## let's Add geam_smooth() what will happen?



```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth()
```
We see there is smooth line appear on the middle of the data points.
## Adding “wiggliness” in the smoothing plot:

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth(span = 0.2) 
```
What changes come in the above graph and previous graph. Let us again check by keeping sapn = 1.

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point()+ geom_smooth(span = 1)
```
We can make sence that by default ggplot kept value of span 1. If we set method= lm inside geom_smooth() we can find stright smooth line. Let us try.

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth(method = lm) 
```
## Let Modify our code little,
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + geom_smooth(method = lm, se= FALSE) 
```
By `se= FALSE` we add smooth line.

## Fixed color

```{r}
ggplot(mpg, aes(displ,hwy)) + geom_point(color = 'red')
```
## Changing color by variable attributes:

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
geom_point()
```
Here we give  colors according to variables name.

## Getting multiple scatterplot of attributes:
We get multiple scatterplot by using `facet_wrap()` function.

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point() + 
facet_wrap(~class)
```
In above figure we find distribution of various variables along with displ and hwy variables.

## Histogram

```{r}
ggplot(mpg, aes(hwy)) + geom_histogram()
```
hwy variable bin automatically.

## Changing bin size of the histogram:

```{r}
ggplot(mpg, aes(hwy)) + geom_histogram(binwidth = 2.5)
```
## Frequency polygon:
A frequency polygon is a line graph of class frequency plotted against class midpoint. It can be obtained by joining the midpoints of the tops of the rectangles in the histogram.

```{r}
ggplot(mpg, aes(hwy)) + geom_freqpoly()
```
## Change Bin size of frequency Polygon

```{r}
ggplot(mpg, aes(hwy)) + geom_freqpoly(binwidth= 1)
```
We can see the effect of binwidth from figure by comparing above figure with previous one.

## Histogram with faceting:
We already discuss about what a facet do in  scatter plot. Similarly in histogram it gives multiple subplots.

```{r}
ggplot(mpg, aes(displ, fill = drv)) + geom_histogram(binwidth = 0.5) + 
facet_wrap(~drv, ncol = 1)
```
## Bar plot


```{r}
ggplot(mpg, aes(manufacturer)) + geom_bar()
```
We can draw bar plot in `geom_bar()` function from bar plot we can see `dodge` and `toyotao` has maximun frequency.

## Let's Use alpha inside geom_ point()


```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point(alpha = 1 / 3)
```
Alpha refers to the opacity of a geom. Values of alpha range from 0 to 1, with lower values corresponding to more transparent colors.

## Mofifying the axes 
```{r}
ggplot(mpg, aes(cty, hwy)) +geom_point(alpha = 1 / 3) + xlab("city driving (mpg)") + 
ylab("highway driving (mpg)")
```




```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point(alpha = 1 / 3) + xlab(NULL) + 
 ylab(NULL)
```






































