---
title: "Pipe"
author: "Durga Pokharel"
date: "02/12/2021"
always_allow_html: true
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(magrittr)
```

# Data Analysis Using Pipe Operator in R

## What is pipe oprator
Pipe operator are a very powerful tool for clearly expresing a sequence of multiple operations. Pipe come from `magrittr` package. Pipe helps us to write code in a way that is easier to read write and understand.


## Why to use Pipe
R is functional language it requeres lots of parenthesis while doing complex programming, this often means we have to nest those parenthesis together for statistical computing. This makes R codes hard to read and understand. In this suitation pipe help us. There are three types of pipe operators. They are,

### %< >% Compound Operator
Updates values for same variable.

```{r}
x <- rnorm(100)
x %<>% abs %>% sort
x
```
Here at first %<>% find absolute values of x and and then it sort in uncreasing order.

### %>% tee Operator

Normally, code ends after plot command but the “tee” pipe operator 
allows it to continue for the next arguement

```{r}
set.seed(123) 
rnorm(200) %>% 
matrix(ncol = 2) %T>% 
plot %>% 
colSums
```


### %$%

This operator is useful when we want to pipe a dataframe which may contains many columns in to a function that is applied to some of the columns.


```{r}
data.frame(z = rnorm(100)) %$%
ts.plot(z)
```


## let's do some Data Analysis in R's Builtin data iris with Pipe Operator:

First, let's us check what attributes `iris` data contains.

```{r}
head(iris)
```
`iris` data have four columns they are `Sepal.Length,Sepal.Width, Petal.Length ,Petal.Width Species`. 


```{r}
library(dplyr)
iris %>%
  select(contains('Sepal')) %>%
  mutate(Sepal.Area=Sepal.Length * Sepal.Width)
```
`contains` function only select those columns which contains **sepal** word. Here we also import `library(dplyr)` dplyr packages is mainly used for data cleaning purpose. It contaims five buitin function they are: 

* select

* filter

* arrange
 
* mutate

These five functions can do majority of data manipulatlation for data analysis task.

In above code mutate function added new column ("Sepal.Area") in the data frame.

```{r}
iris %>%
  select(starts_with('Sepal')) %>%
  filter(Sepal.Length >= 5.0) %>%
  arrange(Sepal.Length)
``` 
By using above code we want to select only those columns which has **Sepal** as prefix and want to filter those values which have values greater than `Sepal.Length >= 5.0` and finally we did arrange in accending order. If we to sort in decending order we simply need to type **desc** inside arrange function as like below.


```{r}
iris %>%
  select(starts_with('Sepal')) %>%
  filter(Sepal.Length >= 5.0) %>%
  arrange(desc(Sepal.Length))
```
## Let's try one Different Step:

Let us observer what will come in result if we use `end_with` in place of `starts_with`.

```{r}
iris %>%
  select(ends_with('Length')) %>%
  rowwise() %>%
  mutate(Length.Diff = Sepal.Length- Petal.Length)

```
I hope we have observed what new came here. Here, we got prefix as **Length**. It means we got only those columns which has length as it's prefix. Also by using mutate functiin we created new column **Length.Diff**.

## Use `transmute` function 


```{r}
iris %>%
  select(contains('Sepal'), Species) %>%
  transmute(Sepal.Area= Sepal.Length * Sepal.Width)
```
Here **transmute** funtion only show new columns it hide Sepal.Length and Sepal.Width columns.

## What a difference we observe when we use pipe operator and genral R function: 
Let us clearify it from code


### Using R general function


Assign value of x with given column vector

```{r}
x <- c(0.109, 0.359, 0.63, 0.996, 0.515, 0.142, 0.017, 0.829, 0.907)
```

Here we want to compute the following with in single line of code,

*  Compute the logarithm of `x`,

*  return suitably lagged and iterated difference

* compute the exponential function and round the result


```{r}
round(exp(diff(log(x))), 1)

```
Let's to some code using pipe operators.


```{r}
x %>%log() %>%
  diff() %>%
  exp()%>%
  round(1)
```
In above pipe come handly because we did not need to nest with lots of parenthesis. Finally we got the same result.

However, there are lots of advantages of pipe operator there are some creteria to use pipe operator. Following are the some of the limitations of pipe operators.

*  Our pipes are longer than (say) ten steps

*  We have multiple inputs or outputs

* We are starting to think about a directed graph with a complex dependency 
structure

* We're doing internal package development

In above mention cases we do not prefer to use pipe operator.
















