---
title: "Project 4"
author: "Durga Pokharel"
date: "05/01/2022"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
# Loading the data here
library(haven)
bank_loan_df <- read_sav("F:/MDS-Private-Study-Materials/First Semester/Statistical Computing with R/Assignments/Data/P4_bankloan_5000_clients.sav")
bank_loan_df$defaulted_loan<-as.factor(bank_loan_df$defaulted_loan)
bank_loan_df$education_level<-as.factor(bank_loan_df$education_level)
str(bank_loan_df)

```
This is a hypothetical data file that concerns a bank's efforts to redu" " ce" "the rate of loan defaults. This file contains financial and demographic" "information on 5000 past customers that the bank will use to create binning rule.

# Train Test Validation
```{r}
library(caret)

```
# Testing the data into training and testing set
```{r}
set.seed(1234)
ind = sample(2,nrow(bank_loan_df),replace = T, prob = c(0.8, 0.3))
train_data <- bank_loan_df[ind==1,]
test_data <- bank_loan_df[ind==2,]
```

# Logistic Regression
## Training the logictic model
```{r,message=FALSE}
logic_model <- train(defaulted_loan~., data = train_data, method = "glm", family= "binomial")
summary(logic_model)
```
# Testing the Logistic model
```{r}
pred1 <- predict(logic_model, test_data)
```

# Confusion Matrix
```{r}
confusionMatrix(pred1, test_data$defaulted_loan)
```
# KNN Model with train/test validation
```{r}
knn_model<-train(defaulted_loan~.,data = train_data,
 method="knn",
 preProcess = c("center", "scale"),
 tuneLength = 10
 )

```

# Obtain the result
```{r}
knn_model$result
```

# Testing the model
```{r}
pred2 <- predict(knn_model, test_data)
```

# Confusion Matrix
```{r}
confusionMatrix(pred2,test_data$defaulted_loan)
```
# Fitting Naive Bayes 
## Training the model
```{r}
library(e1071)
naive_model <- naiveBayes(defaulted_loan~., train_data)
summary(naive_model)
```
## Testing the model
```{r}
pred3 <- predict(naive_model, test_data)


```

## Confusion Matrix
```{r}
confusionMatrix(pred3,test_data$defaulted_loan)
```
# Support Vector Machine (SVM) Model
## Training the model
```{r}
svm_model <- svm(formula= defaulted_loan~., data = train_data, type = "C-classification", kernel= "linear")
summary(svm_model)
```
## Testing the Model
```{r}
pred4 <- predict(svm_model, test_data)

```

## Confusion Matrix
```{r}
confusionMatrix(pred4, test_data$defaulted_loan)
```
# Decision Tree Model
## Training the model
```{r}
decision_tree_model <-train(defaulted_loan~.,
 data = train_data,
 method="rpart",
 parms = list(split = "information"),
 tuneLength=10
 )


```
## Testing the model
```{r}
pred5 <- predict(decision_tree_model, test_data)

```
## Confusion Matrix
```{r}
confusionMatrix(pred5, test_data$defaulted_loan)
```
# Artifical Neural Network (ANN) Model
## Training the Model
```{r}
ann_model <- train(defaulted_loan ~ ., data = train_data, 
method = "nnet",
preProcess = c("center","scale"), 
maxit = 250, # Maximum number of iterations
tuneGrid = data.frame(size = 1, decay = 0),
 # tuneGrid = data.frame(size = 0, decay = 0),skip=TRUE, # Technically, this is log-reg
metric = "Accuracy")


```
## Testing the model
```{r}
pred6 <- predict(ann_model, test_data)
```

## Confusion Matrix
```{r}
confusionMatrix(pred6, test_data$defaulted_loan)
```
# Leave one Out Validation
## Read the data



```{r}
library(haven)
bank_loan_df <- read_sav("F:/MDS-Private-Study-Materials/First Semester/Statistical Computing with R/Assignments/Data/P4_bankloan_5000_clients.sav")
bank_loan_df$defaulted_loan<-as.factor(bank_loan_df$defaulted_loan)
bank_loan_df$education_level<-as.factor(bank_loan_df$education_level)

```

# Logistic Regression With LOOCV Validation
## Training Logistic Regression Model
```{r}
set.seed(1234)
library(caret)

```

```{r}
ind<-sample(2,nrow(bank_loan_df),replace=T,prob = c(0.7,0.3))
 train_data<-bank_loan_df[ind==1,]
 test_data<-bank_loan_df[ind==2,]

```

## Setting Up the Train Control
```{r}
loocv_train_control<-trainControl(method = "LOOCV")
```
# Logistic Regression With LOOCV Validation
## Training Logistic Regression Model
```{r}
#logistic_clf1<-train(defaulted_loan~.,
 #data=train_data,
 #method="glm",
 #family="binomial",
 #trControl=loocv_train_control, 
 #verbose=F
# )


```
# KNN Model with LOOCV validation
## Training KNN Model
```{r}
knn_clf1<-train(defaulted_loan~.,data = train_data,
 method="knn",
 trControl=loocv_train_control
 )


```
## Obtain the result
```{r}
knn_clf1$result
```
# Confusion Matrix for Model Evaluation
```{r}
predicted_val_knn1<-predict(knn_clf1,newdata = test_data)
confusionMatrix(predicted_val_knn1,test_data$defaulted_loan)

```
# Naïve Bayes classifier
## Training the Model
```{r}
nb_clf1<-train(defaulted_loan~.,
 data=train_data,
 method="naive_bayes",
 usepoisson = TRUE,
 trControl=loocv_train_control
 )


```
# Making Prediction on Test Data
```{r}
predicted_val_nb1<-predict(nb_clf1,newdata = test_data)

```
# Confusion Matrix for Model Evaluation
```{r}
confusionMatrix(predicted_val_nb1,test_data$defaulted_loan)
```
# Support Vector Machine (SVM) Model
```{r}
#ctrl <- trainControl(method = "LOOCV", savePred=T)
 #svm_clf1<-train(defaulted_loan~.,
 # data=train_data,
 # method="svmLinear",
 # trControl=ctrl,
 # )
 #svm_clf

```

# Making the Prediction for test data

```{r}
#predicted_val_svm1<-predict(svm_clf1,newdata = test_data)
```

# Confusion Matrix for Model Evaluation
```{r}
#confusionMatrix(predicted_val_svm1,test_data$defaulted_loan)
```
 The Model did not Converge to a solution. Leaving it as is for now.
# Decision Tree Model
```{r}
#dtree_clf1<-train(defaulted_loan~.,
#data = train_data,
#method="rpart",
#parms = list(split = "information"),
#tuneLength=10,
# trControl=loocv_train_control
 #)
#dtree_clf1

```

# Making the Prediction for test data
```{r}
#predicted_val_dtree1<-predict(dtree_clf1,newdata = test_data)
```

# Confusion Matrix for Model Evaluation
```{r}
# confusionMatrix(predicted_val_dtree1,test_data$defaulted_loan)
```

# Artifical Neural Network (ANN) Model
## Training the Model
```{r}
#ann_clf1 <- train(defaulted_loan ~ ., data = train_data, 
# method = "nnet",
 # preProcess = c("center","scale"), 
# maxit = 250, # Maximum number of iterations
 # tuneGrid = data.frame(size = 1, decay = 0),
 # tuneGrid = data.frame(size = 0, decay = 0),skip=TRUE, # Technically, this is log-reg
 # metric = "Accuracy",
 # trControl=loocv_train_control)


```

## Making the Predictions for Test data
```{r}
#predicted_val_ann1<-predict(ann_clf1,newdata = test_data)
```

## Confusion Matrix for the Model Evaluation
```{r}
#confusionMatrix(predicted_val_ann1,test_data$defaulted_loan)
```

# K-Fold Cross Validation
## Reading the File
```{r}
library(haven)
bank_loan_df <- read_sav("F:/MDS-Private-Study-Materials/First Semester/Statistical Computing with R/Assignments/Data/P4_bankloan_5000_clients.sav")

```

## Changing the data type of variables
```{r}
bank_loan_df$defaulted_loan<-as.factor(bank_loan_df$defaulted_loan)
bank_loan_df$education_level<-as.factor(bank_loan_df$education_level)

```

## Splitting the data into train and test set
```{r}
set.seed(1234)
library(caret)
ind<-sample(2,nrow(bank_loan_df),replace=T,prob = c(0.7,0.3))
train_data<-bank_loan_df[ind==1,]
test_data<-bank_loan_df[ind==2,]


```

## Setting Up the Train Control
```{r}
cv_train_control<-trainControl(method = "cv",number = 10)
```

## Logistic Regression With Cross Validation
## Training Logistic Regression Model
```{r}
logistic_clf1<-train(defaulted_loan~.,
 data=train_data,
 method="glm",
 family="binomial",
 trControl=cv_train_control
 )
 summary(logistic_clf1)


```
## Making the Prediction
```{r}
predicted_val_log1<-predict(logistic_clf1,newdata = test_data)
```

## Confusion Matrix for Evaluation
```{r}
confusionMatrix(predicted_val_log1,test_data$defaulted_loan)
```
## KNN Model with Cross validation
## Training KNN Model
```{r}
knn_clf1<-train(defaulted_loan~.,data = train_data,
 method="knn",
 trControl=cv_train_control
 )

```

## Getting the Result of the Model
```{r}
knn_clf1$result
```

## Confusion Matrix for Model Evaluation
```{r}
predicted_val_knn1<-predict(knn_clf1,newdata = test_data)
confusionMatrix(predicted_val_knn1,test_data$defaulted_loan)

```
# Naïve Bayes classifier
## Training the Model
```{r}
library(naivebayes)
```
```{r}
nb_clf1<-train(defaulted_loan~.,
 data=train_data,
 method="naive_bayes",
 usepoisson = TRUE,
 trControl=cv_train_control
 )
summary(nb_clf1)

```
##  Making Prediction on Test Data
```{r}
predicted_val_nb1<-predict(nb_clf1,newdata = test_data)
```

## Confusion Matrix for Model Evaluation
```{r}
confusionMatrix(predicted_val_nb1,test_data$defaulted_loan)
```
# Support Vector Machine (SVM) Model
## Training the Model
```{r}
#svm_clf1<-train(defaulted_loan~.,
#data=train_data,
#method="svmLinear",
#trControl=cv_train_control,
 #)
#svm_clf1

```

# Making the Prediction for test data
```{r}
# predicted_val_svm1<-predict(svm_clf1,newdata = test_data)
```

# Confusion Matrix for Model Evaluation
```{r}
# confusionMatrix(predicted_val_svm1,test_data$defaulted_loan)
```
# Decision Tree Model
```{r}
#dtree_clf1<-train(defaulted_loan~.,
#data = train_data,
#method="rpart",
#parms = list(split = "information"),
#tuneLength=10,
#trControl=cv_train_control
 #)
 #dtree_clf1

```

# Making the Prediction for test data
```{r}
#predicted_val_dtree1<-predict(dtree_clf1,newdata = test_data)
```
# Confusion Matrix for Model Evaluation
```{r}
#confusionMatrix(predicted_val_dtree1,test_data$defaulted_loan)
```

# Artifical Neural Network (ANN) Model
## Training the Model
```{r}
#ann_clf1 <- train(defaulted_loan ~ ., data = train_data, 
#method = "nnet",
# preProcess = c("center","scale"), 
#maxit = 250, # Maximum number of iterations
 #tuneGrid = data.frame(size = 1, decay = 0),
 # tuneGrid = data.frame(size = 0, decay = 0),skip=TRUE, # Technically, this is log-reg
#metric = "Accuracy",
 #trControl=cv_train_control)

```

# Making the Predictions for Test data

```{r}
# predicted_val_ann1<-predict(ann_clf1,newdata = test_data)
```

# Confusion Matrix for the Model Evaluation
```{r}
# confusionMatrix(predicted_val_ann1,test_data$defaulted_loan)
```

# Repeated K-Fold Cross Validation
# Setting Up the Train Control
```{r}
#rep_cv_train_control<-trainControl(method = "repeatedcv",number = 10,repeats = 3)
```
## Logistic Regression With Repeated Cross Validation
## Training Logistic Regression Model
```{r}
#logistic_clf1<-train(defaulted_loan~.,
# data=train_data,
# method="glm",
 #family="binomial",
 #trControl=rep_cv_train_control
 #)
 #summary(logistic_clf1)

```

## Making the Prediction
```{r}
#predicted_val_log1<-predict(logistic_clf1,newdata = test_data)
```

## Confusion Matrix for Evaluation
```{r}
#confusionMatrix(predicted_val_log1,test_data$defaulted_loan)
```
## KNN Model with Repeated Cross validation
## Training KNN Model
```{r}
#knn_clf1<-train(defaulted_loan~.,data = train_data,
# method="knn",
 #trControl=rep_cv_train_control
 #)

```

## Getting the Result of the Model
```{r}
#knn_clf1$result
```

## Confusion Matrix for Model Evaluation
```{r}
#predicted_val_knn1<-predict(knn_clf1,newdata = test_data)
#confusionMatrix(predicted_val_knn1,test_data$defaulted_loan)

```

## Naïve Bayes classifier
## Training the Model
```{r}
#library(naivebayes)
```
```{r}
#nb_clf1<-train(defaulted_loan~.,
 #data=train_data,
 #method="naive_bayes",
 #usepoisson = TRUE,
 #trControl=rep_cv_train_control
 #)
#summary(nb_clf1)

```

## Making Prediction on Test Data
```{r}
#predicted_val_nb1<-predict(nb_clf1,newdata = test_data)

```

## Confusion Matrix for Model Evaluation

```{r}
#confusionMatrix(predicted_val_nb1,test_data$defaulted_loan)
```

## Support Vector Machine (SVM) Model
## Training the model
```{r}
#svm_clf1<-train(defaulted_loan~.,
 #data=train_data,
 #method="svmLinear",
 #trControl=rep_cv_train_control,
 #)
#svm_clf1

```
## Making the Prediction for test data
```{r}
#predicted_val_svm1<-predict(svm_clf1,newdata = test_data)
```

## Confusion Matrix for Model Evaluation
```{r}
#confusionMatrix(predicted_val_svm1,test_data$defaulted_loan)
```
## Decision Tree Model
```{r}
#dtree_clf1<-train(defaulted_loan~.,
 #data = train_data,
 #method="rpart",
 #parms = list(split = "information"),
 #tuneLength=10,
 #trControl=rep_cv_train_control
 #)
#dtree_clf1

```

## Making the Prediction for test data

```{r}
 #predicted_val_dtree1<-predict(dtree_clf1,newdata = test_data)
```
## Confusion Matrix for Model Evaluation

```{r}
# confusionMatrix(predicted_val_dtree1,test_data$defaulted_loan)
```

## Artifical Neural Network (ANN) Model
## Training the Model
```{r}
#ann_clf1 <- train(defaulted_loan ~ ., data = train_data, 
#method = "nnet",
#preProcess = c("center","scale"), 
#maxit = 250, # Maximum number of iterations
# tuneGrid = data.frame(size = 1, decay = 0),
 # tuneGrid = data.frame(size = 0, decay = 0),skip=TRUE, # Technically, this is log-reg
 #metric = "Accuracy",
 #trControl=rep_cv_train_control)

```
## Making the Predictions for Test data
```{r}
#predicted_val_ann1<-predict(ann_clf1,newdata = test_data)
```

## Confusion Matrix for the Model Evaluation
```{r}
#confusionMatrix(predicted_val_ann1,test_data$defaulted_loan)
```

## Bagging, Boosting and Random Forest
## Reading the File
```{r}
library(haven)
bank_loan_df <- read_sav("F:/MDS-Private-Study-Materials/First Semester/Statistical Computing with R/Assignments/Data/P4_bankloan_5000_clients.sav")
bank_loan_df$defaulted_loan<-as.factor(bank_loan_df$defaulted_loan)
bank_loan_df$education_level<-as.factor(bank_loan_df$education_level)

```

## Splitting the data into train and test set
```{r}
set.seed(1234)
 library(caret)
## Loading required package: ggplot2
## Loading required package: lattice
ind<-sample(2,nrow(bank_loan_df),replace=T,prob = c(0.7,0.3))
 train_data<-bank_loan_df[ind==1,]
 test_data<-bank_loan_df[ind==2,]

```
## Bagging Model
## Training the Model
```{r}
library("ipred")
 bag_dtree_clf<-bagging(defaulted_loan~.,
 data = train_data,
 coob=T
 )
print(bag_dtree_clf)

```
## Making the Prediction
```{r}
predicted_bag_tree<-predict(bag_dtree_clf,newdata = test_data)
library(caret)
 confusionMatrix(predicted_bag_tree,test_data$defaulted_loan)


```
## Random Forest Model
## Training the Model
```{r}
set.seed(1234)
library(randomForest)
rf_clf<-randomForest(defaulted_loan~.,
 data = train_data)
rf_clf



```
## Making the Prediction

```{r}
predicted_rf<-predict(rf_clf,newdata = test_data)
confusionMatrix(predicted_rf,test_data$defaulted_loan)


```
## Extreme Gradient Boosting
## Training the Model
```{r}
#xglm_clf<-train(defaulted_loan~.,
 #data = train_data,
 #method="xgbTree",
 #verbose=F
# )


```
## Making the Prediction
```{r}
#predicted_xgb<-predict(xglm_clf,newdata = test_data)
#confusionMatrix(predicted_xgb,test_data$defaulted_loan)


```


